---
title: "Application: Cross-sample Comparison"
author: "Ruiqi Li"
date: "08/20/2024"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Application: Cross-sample Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Loading Data
One of the downstream tasks for LMD is cross-sample comparison to identify sample-enriched or shared cell populations without the need for integration and differential abundance testing. Here we use a 10x mouse embryonic skin dataset. The processing step can be found [here](https://github.com/KlugerLab/LocalizedMarkerDetector/blob/main/LMD-paper/analysis/mouse_smoM2_example_process_from_10x.R). The 10x data and processed Seurat data are available from the corresponding authors upon reasonable request.

```{r, warning = FALSE, message=FALSE}
##### Load required R libraries
library("LocalizedMarkerDetector")
library("ggplot2")
library("Seurat")
library("SeuratWrappers")
library("patchwork")
```

This dataset includes a SmoM2 mutant sample with a paired wildtype sample, both collected at embryonic day 13.5. We can visualize these two samples using UMAP, with cells color-coded by cell types.
```{r, warning = FALSE, message=FALSE}
##### Load Data
dir.path0 = "~/" # Specify the directory path where the data is saved
```
```{r, warning = FALSE, message=FALSE, include = FALSE}
source("paths.R")
```
```{r load_data, fig.height=3, fig.width=6, warning = FALSE, message=FALSE}
data_S_MUT <- readRDS(file.path(dir.path0,"smom2/process_data/data_S_smom2_dermal_E13.5_MUT.rds"))
data_S_CTL <- readRDS(file.path(dir.path0,"smom2/process_data/data_S_smom2_dermal_E13.5_CTL.rds"))

p1 = DimPlot(data_S_MUT, 
          group.by = "celltype", shuffle = TRUE, cols = c("DC"="#d1495b", "UD" = "#edae49", "LD" = "#d3d3d3")) + ggtitle("Mutant") & NoAxes()
p2 = DimPlot(data_S_CTL, 
          group.by = "celltype", shuffle = TRUE, cols = c("DC"="#d1495b", "UD" = "#edae49", "LD" = "#d3d3d3")) + ggtitle("Wildtype") & NoAxes()
p1 + p2 + plot_layout(guides = "collect")
```

# Running LMD and Identifying localized gene modules

Next, we run LMD and identify gene modules on the mutant sample.
```{r, warning = FALSE, message=FALSE}
#### Prepare data
dat = as.matrix(data_S_MUT[["RNA"]]@data)
# Keep genes express > 10 cells and less than 50% of cells
Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = names(Gene_detected_count)[(Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)]
dat = dat[selected_genes,,drop = FALSE]
# Extract top 14 PCs (# of PC based on Elbow Plot)
feature_space = Embeddings(data_S_MUT[["pca"]])[,1:FindPC(data_S_MUT)]
```
```{r, warning = FALSE, message=FALSE}
##### RunLMD
lmd_score_res = LMD(dat, feature_space, max_time = 2^20)

res = show_result_lmd(lmd_score_res)
```
```{r, warning = FALSE, message=FALSE}
##### Identify Gene Modules
if(!"alra" %in% names(data_S_MUT@assays)){
  data_S_MUT <- RunALRA(data_S_MUT, assay = "RNA")
  DefaultAssay(data_S_MUT) <- "RNA"
}
dat_alra = as.matrix(data_S_MUT[["alra"]]@data)[names(res$cut_off_gene),]
dist = CalculateGeneDistance(dat_alra, method = "jaccard")

gene_partition = ClusterGenes(dist, clustering_method = "average", deepSplit = 2)
levels(gene_partition) = 1:nlevels(gene_partition)
table(gene_partition)
```

# Computing module activity scores for both mutant and wildtype samples

Next, we calculate the module activity score for each module in both the mutant and wildtype samples.

**Note**: This step is CPU-intensive and may take 4 to 5 minutes to complete. 

```{r, warning = FALSE, message=FALSE}
data_S_MUT = AddModuleActivityScore(data_S_MUT, gene_partition = gene_partition)
data_S_CTL = AddModuleActivityScore(data_S_CTL, gene_partition = gene_partition)
```

Then we can visualize the expression patterns of gene modules across both samples.

## Modules reflecting functional similar cell populations between mutant and WT

Here we show two modules that are localized in both mutant and wildtype. 

```{r shared_module, fig.height=6, fig.width=7, warning = FALSE, message=FALSE}
library(ggplotify)
pl = lapply(c("Module1","Module16"), function(module_name){
  pl = lapply(c("MUT","CTL"), function(i){
    FeaturePlot(get(paste0("data_S_",i)), features = c(module_name), order = TRUE) + scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + ggtitle(i) & labs(color = "ModuleScore") & NoAxes()
  })
  p = as.ggplot(wrap_plots(pl, nrow = 1) + plot_layout(guides = "collect") + plot_annotation(title = module_name))
})
p = wrap_plots(pl,ncol = 1)
p
```

## Modules reflecting mutant-enriched cell populations

Here we show two modules that are only localized in mutant and dispersed in wildtype.

```{r specific_module, fig.height=6, fig.width=7, warning = FALSE, message=FALSE}
library(ggplotify)
pl = lapply(c("Module14","Module17"), function(module_name){
  pl = lapply(c("MUT","CTL"), function(i){
    FeaturePlot(get(paste0("data_S_",i)), features = c(module_name), order = TRUE) + scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + ggtitle(i) & labs(color = "ModuleScore") & NoAxes()
  })
  p = as.ggplot(wrap_plots(pl, nrow = 1) + plot_layout(guides = "collect") + plot_annotation(title = module_name))
})
p = wrap_plots(pl,ncol = 1)
p
```

